---
title: "p8105_hw1_yc4804"
output: html_document
date: "2025-09-20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1

```{r}

#load data packages
library(moderndive)
library(ggplot2)
data("early_january_weather")


```
The dataset early_january_weather contains hourly weather observations at LaGuardia Airport in early January 2013. 
The variable names are: `r paste(names(early_january_weather), collapse = ", ")`.
Important variables for our analysis are time_hour, temp, and humid.
The dataset has `r nrow(early_january_weather)` rows and `r ncol(early_january_weather)` columns.
 The average air temperature is `r round(mean(early_january_weather$temp, na.rm = TRUE), 1)` °F.

```{r}

p_temp_time<-ggplot(
  early_january_weather,
  aes(x = time_hour, y = temp, color = humid)
) +
  geom_point(alpha = 0.6) +
  labs(
    x = "Time (Hourly)",
    y = "Temperature (°F)",
    color = "Humidity",
    title = "NYC Early January Weather: Temperature vs Time"
  ) 
p_temp_time

```


Temperatures show clear diurnal cycles and a gradual warm-up into mid-January, peaking around Jan 13–14. Humidity (color scale) tends to be higher at higher temperatures—lighter points are concentrated during the warm spell—while darker (drier) points are more common during cooler periods and overnight. Overall, the figure suggests a modest positive temp–humidity relationship during this window, with day–night fluctuations superimposed on the warming trend.

